---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# multibridge: Evaluating Multinomial Order Restrictions with Bridge Sampling

<!-- badges: start -->
[![CRAN/METACRAN](https://img.shields.io/cran/v/multibridge?label=CRAN&logo=r)](https://cran.r-project.org/web/packages/multibridge/index.html) [![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active) ![GitHub last commit (devel)](https://img.shields.io/github/last-commit/ASarafoglou/multibridge/devel?label=Last commit&logo=github&logoColor=%23FFF) ![Travis build status](https://img.shields.io/travis/ASarafoglou/multibridge?label=Build&logo=travis-ci&logoColor=%23FFF) [![codecov](https://codecov.io/gh/ASarafoglou/multibridge/branch/master/graph/badge.svg)](https://codecov.io/gh/ASarafoglou/multibridge) [![GitHub bug issues](https://img.shields.io/github/issues/ASarafoglou/multibridge/bug?label=Bugs&logo=github&logoColor=%23FFF)](https://github.com/ASarafoglou/multibridge/issues?q=is%3Aopen+is%3Aissue+label%3Abug)
<!-- badges: end -->

Evaluates hypotheses concerning the distribution of multinomial proportions using bridge sampling.
The bridge sampling routine is able to compute Bayes factors for hypotheses that entail inequality constraints, equality constraints, free parameters, and a mix of all three.
These hypotheses are tested against the encompassing hypothesis, that all parameters vary freely.

## Installation

You can install the released version of `multibridge` from [CRAN](https://CRAN.R-project.org) with:

~~~r
install.packages("multibridge")
~~~

And the development version from [GitHub](https://github.com/) with:

~~~r
# install.packages("remotes")
remotes::install_github("ASarafoglou/multibridge")
~~~


## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library("multibridge")
## basic example code
```


## Package dependencies

```{r echo = FALSE, message = FALSE, warning = FALSE}
library("depgraph")

plot_dependency_graph()
```
