---
title: "Memory of Lifestresses"
author: "Alexandra Sarafoglou"
date: "9/16/2020"
output: html_document
vignette: >
  %\VignetteIndexEntry{Memory of Lifestresses}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
bibliography: ../inst/REFERENCES.bib
---

In this vignette, we will explain how to compute a Bayes factor for inequality-constrained hypotheses for multinomial models. 

## Model and Data

As example for an inequality-constrained hypothesis in multinomial models, we will use the data set, `lifestresses`, which is included in the package `multibridge`. The dataset is based on a survey study by @uhlenhuth1974symptom which among other things has dealt with experienced negative life events and lifestresses. The subset of the data which we use here summarizes when respondents have experienced a negative event in their lives in the 18 months prior to the interview. In total, the dataset contains the responses of 147 people. This subset was analyzed by @haberman1978 who wanted to show that the participants listed mainly negative events of the recent past. Furthermore, in the context of inequality-constrained hypotheses the dataset was discussed in @sarafoglou2020evaluatingPreprint. 

To get an overview, we will load and visualize the data before we start the data analysis.

```{r}
library(multibridge)
data(lifestresses)
lifestresses

# visualize the data
op <- par(cex.main = 1.5, mar = c(5, 6, 4, 5) + 0.1, mgp = c(3.5, 1, 0), cex.lab = 1.5 , font.lab = 2, cex.axis = 1.3, bty = "n", las = 1)
  plot(lifestresses$month, lifestresses$stress.freq, col = "black", 
       pch = 1, cex = 2, type="b",
       xlim = c(0, 18), ylim = c(0, 20), 
       ylab = "", xlab = "", axes = FALSE)
  axis(1, at = seq(0, 18, length.out = 10), label = seq(0, 18, length.out = 10))
  axis(2, at = seq(0, 20, length.out = 5), label = seq(0, 20, length.out = 5), las = 1) 
  par(las = 0)
  mtext("Months Before Interview", side = 1, line = 2.5, cex = 1.5)
  mtext("Participants Reporting a\nNegative Life Event", side = 2, line = 3.0, cex = 1.5)
```

The model that we will use assumes that the vector of observations $x_1, \cdots, x_K$ in the $K$ categories follow a multinomial distribution. The parameter vector of the multinomial model, $\theta_1, \cdots, \theta_K$, contains the probabilities of observing a value in a particular category; here, it reflects the probabilities of reporting a negative life event or life stress during one of the 18 months. The parameter vector $\theta_1, \cdots, \theta_K$ is drawn from a Dirichlet distribution with concentration parameters $\alpha_1, \cdots, \alpha_K$. The model can be described as follows:

\begin{align}
  x_1, \cdots, x_K &\sim \text{Multinomial}(\sum_{k = 1}^K x_k, \theta_1, \cdots, \theta_K) \\
  \theta_1, \cdots, \theta_K &\sim \text{Dirichlet}(\alpha_1, \cdots, \alpha_K) \\
\end{align}

Here, we test the inequality-constrained hypothesis $\mathcal{H}_r$ that the number of reported negative life events decreases over time against the encompassing hypothesis $\mathcal{H}_e$ without constraints:

\begin{align*}
    \mathcal{H}_r &: \theta_{1} > \theta_{2} > \cdots > \theta_{18} \\
    \mathcal{H}_e &: \theta_1, \theta_2,  \cdots, \theta_{18}.
\end{align*}

To compute the Bayes factor in favor of the inequality-constrained hypothesis, $\text{BF}_{re}$, we need to specify
(1) and a vector containing the number of observations, (2) the inequality-constrained hypothesis,
(3) a vector with concentration parameters, and (4) the categories of interest 
(i.e., months prior to the interview).   . 

```{r}
x <- lifestresses$stress.freq
# Prior specification 
# We assign a uniform Dirichlet distribution, that is, we set all concentration parameters to 1
a <- rep(1, 18)
# Test the following restricted Hypothesis:
# Hr: month1 > month2 > ... > month18 
Hr  <- paste0(1:18, collapse=">"); Hr
categories <- lifestresses$month
```

With this information, we can now conduct the analysis with the function `multBfInformed()`. 
Since we are interested in quantifying evidence in favor of the restricted hypothesis, we 
set the Bayes factor type to `BFre`. For reproducibility, we are also setting a 
seed with the argument `seed`:

```{r, cache = TRUE}
ineq_results <- multBfInformed(x=x, Hr=Hr, a=a, factor_levels=categories,
                                 bf_type = 'BFre', seed = 2020)
```

## Summarize the Results

We can get a quick overview of the results by using the implemented `summary()` method:

```{r}
summary(ineq_results)
```

Since we are dealing with many categories, one might want to assess the accuracy
of the bridge sampling estimate. This can be done, for instance, using the `S3` 
method `bridge_output`. Using this method, we can see, for instance coefficient of 
variation (with respect to the marginal likelihood) expressed as a percentage
(see @fruhwirth2004estimating and  @gronau2017tutorial for details).

```{r}
bridge_info <- bridge_output(ineq_results)
bridge_info
```

The percentage error is about `r bridge_info[[1]]$post$error_measures$percentage`,
which can be considered low. More information on the accuracy can be extracted
from the `bridge_info` list. For instance, we could extract the full list of
error measures, which features the approximate relative mean-squared error for the 
marginal likelihood estimate, the approximate coefficient of variation for the marginal 
likelihood estimate (assumes that bridge estimate is unbiased), and 
the approximate percentage error of the marginal likelihood estimate.

```{r}
bridge_info[[1]]$post$error_measures
```

Note that the `bridge_output` only gives us the 
bridge sampling estimate a associated with the the log marginal likelihood estimate 
of the posterior distribution. This is due
to the fact that we assigned a uniform distribution as prior to the model parameters.
In these cases the bridge sampling algorithm is not necessary since log marginal
likelihood estimate of the prior distribution can be computed analytically.

# References